---
title: "EDA-III"
output: rmarkdown::github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  error = FALSE,
  eval = FALSE,
  echo = TRUE,
  comment = "#>",
  fig.align='center'
)
```

# Introduction

S aureus and S intermedius are coagulase positive. All other staphylococci are coagulase negative (ncbi.nlm.nih.gov/books/NBK8448).

[Wiki COAG(+)](https://en.wikipedia.org/wiki/Coagulase): Staphylococcus aureus subsp. anaerobius, S. aureus subsp. aureus, S. delphini, S. hyicus, S. intermedius, S. lutrae, and Staphylococcus schleiferi subsp. coagulans.

[Wiki COAG(-)](https://en.wikipedia.org/wiki/Coagulase): S. saprophyticus, S.cohnii subsp. cohnii, S. cohnii subsp. urealyticum, S. captitus subsp. captitus, S. warneri, S.hominis, S.epidermidis, S. caprae, and S.lugdunensis

```{r setup}
devtools::install_github('chebuu/StaphAureus')

library(StaphAureus)

## Imports:
# library(gt)
# library(grid)
# library(dplyr)
# library(ggplot2)
# library(cowplot)
# library(gridExtra)
# library(kableExtra)
```

```{r internal, eval=T, include=F}
library(StaphAureus)

# http://www.sthda.com/english/wiki/wiki.php?id_contents=7930
# https://blog.rstudio.com/2020/04/08/great-looking-tables-gt-0-2/
# https://cran.r-project.org/web/packages/gridExtra/vignettes/tableGrob.html

library(gt)
library(grid)
library(dplyr)
library(ggplot2)
library(cowplot)
library(gridExtra)
library(kableExtra)

# usethis::use_package('gt')
# usethis::use_package('grid')
# usethis::use_package('dplyr')
# usethis::use_package('ggplot2')
# usethis::use_package('cowplot')
# usethis::use_package('gridExtra')
# usethis::use_package('kableExtra')
```

```{r plot_setup, eval=T, include=F}
ggplot2::theme_update(
  plot.title.position = 'plot',
  plot.caption.position = 'plot',
  plot.title = element_text(hjust = 0.5),
  plot.caption = element_text(hjust = 0.5),
  plot.subtitle = element_text(hjust = 0.5)
) %+replace% theme_bw()
```

### Study Population
```{sql}
drop materialized view if exists orgids cascade;
create materialized view orgids as
select distinct on (org_itemid), org_name, org_itemid
from microbiologyevents 
where org_name not ilike '%staph%'
```
```{sql}
drop materialized view if exists staphids cascade;
create materialized view staphids as
select distinct on (org_itemid), org_name, org_itemid
from microbiologyevents 
where org_name ilike '%staph%'
```
```{r staphids, echo=F, eval=T, include=T}
data(staphids)
kable(staphids)
```

```{sql}
-- Confirmed positive and negative tests
drop materialized view if exists study_population cascade;
create materialized view study_population as 
select
	pt.subject_id, 
	mb.hadm_id, 
	pt.expire_flag, 
	mb.org_name, 
	mb.isolate_num, 
	mb.interpretation 
from microbiologyevents mb
left join patients pt 
  on mb.subject_id = pt.subject_id
  and mb.org_name ilike '%STAPH%AUREUS%'
group by 
  pt.subject_id, 
  mb.hadm_id, 
  mb.org_name, 
  mb.isolate_num, 
  mb.interpretation,
  pt.expire_flag
```

Interpretation levels:

`"None"`
  - Negative test result (NULL value)
`"R"`
  - Positive test result (resistant)
`"S"`
  - Positive test result (susceptible)
`"I"`
  - Positive test result (intermediate)
`"P"`
  - Positive test result (S/R not tested)

#### Negative Control 

All NULL `population` test interpretation values.

```{sql}
-- Confirmed negative subset
drop materialized view if exists control cascade;
create materialized view control as
select * from study_population where interpretation is NULL
```

#### Positive Cohort COAG(+) *S. aureus*

All non-NULL `population` test interpretation values.

```{sql}
-- Confirmed positive subset
drop materialized view if exists cohort cascade;
create materialized view cohort as
select * from study_population where interpretation is not NULL
```

```{r cohort_iso, echo=T, eval=T, fig.width=5}
data(cohort)

(
  cohort.hist.isolates <<- {
    cohort.iso.table <<- cohort %>%
    distinct_at(
      vars(subject_id, org_name, isolate_num)
    ) %>%
    mutate(
      org_name = case_when(
        grepl('NEG', org_name) ~ 'COAG (-)',
        grepl('+', org_name)   ~ 'COAG (+)'
      )
    )
  } %>% 
  ggplot(aes(x=isolate_num)) +
    geom_histogram(stat='count') +
    facet_grid(~org_name) +
    labs(
      title = 'Staph isolates by coagulase',
      subtitle = '(confirmed positive)'
    ) +
    theme_bw() + theme(
      plot.title.position = 'plot',
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
    ) 
)
```

```{r cohort_mrt, echo=T, eval=T, fig.width=5}
(
  cohort.hist.mortality <- {
    cohort.mort.table <<- cohort %>%
    group_by(
      subject_id, org_name, isolate_num, expire_flag
    ) %>%
    mutate(
      org_name = case_when(
        grepl('NEG', org_name) ~ 'COAG (-)',
        grepl('+', org_name)   ~ 'COAG (+)'
      )
    )
  } %>% 
  ggplot(aes(x=isolate_num)) +
    geom_histogram(stat='count') +
    facet_grid(~org_name + expire_flag) +
    labs(
      title = 'Patient mortality by coagulase/isolate',
      subtitle = '(confirmed positive)'
    ) + 
    theme_bw() + theme(
      plot.title.position = 'plot',
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
    ) 
)
```

```{r cohort_tab, eval=T, echo=T, fig.align='right'}
cohort.iso.table -> cohort.table

.displayP <- cohort.table %>% filter(grepl('\\+', org_name))
gt(.displayP %>% head(6)) %>%
  fmt_number(columns = vars(isolate_num), decimals = 0) %>%
  fmt_passthrough(columns = vars(subject_id, org_name)) %>%
  tab_header(
    title = md('COAG (+)'),
    subtitle = sprintf('(N=%s)', nrow(.displayP))
  ) %>%
  tab_source_note(md('Subsample of cohort (confirmed positive)'))

.displayN <- cohort.table %>% filter(grepl('-', org_name)) 
gt(.displayN %>% head(6)) %>%
  fmt_number(columns = vars(isolate_num), decimals = 0) %>%
  fmt_passthrough(columns = vars(subject_id, org_name)) %>%
  tab_header(
    title = md('COAG (-)'),
    subtitle = sprintf('(N=%s)', nrow(.displayN))
  ) %>%
  tab_source_note(md('Subsample of cohort (confirmed positive)'))
```
